Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-11-21 paper/main.tex
reference figure in bullet points
move text upfron to methods section and discuss in more detail how number of parameters grows

\textbf{TODO Talk about how to fit individual bacteria and why it is important.}

@@

We have already suspected (see Table~\ref{table:simulation-aspects}) that in order to properly
estimate parameters of our system, we need to account for natural cell-variability, thus assigning
parameters on a per-cell basis.
We have previously presented a list of $N_p$ parameters (see Table~\ref{table:list-of-parameters})
which will either be assigned individually or globally.
We filter out a subset $N_g\leq N_p$ of global parameters, treating the remaining ones individually.
Given $N_c$ cells, the total amount of parameters to be estimated is:
\begin{equation}
    N_{p,\text{tot}} = N_g + N_c (N_p - N_g).
\end{equation}
It is clear, that we should avoid, to needlessly treat parameters non-globally since this increases
the size of the parameter space considerably.
The volume that needs to be sampled to optimize within a given parameter space scales exponentially
with the dimension ie. the total number of parameters $N_{p,\text{tot}}$.
\textbf{[Das ist leider komplett unverständlich. Ich glaube, ich weiss schon, was du sagen möchtest, aber das muss umformuliert werden. Stichworte sind hier: Cell Variability, Heterogeneity, latent variables. Parameter per cell and individual parameters versteht auch niemand den Unterschied.]}\\
\textit{Ich habe versucht, es klarer zu machen. Bitte dazu Feedback!}\\
\textbf{Es muss noch im Methoden Teil mehr erwähnt werden.}

@@

\begin{itemize}
    \item Quadratic Scaling comes from more dense simulation
    \item Overall number of agents which can be simulated is satisfactory for current purposes
    \item Improvement from $1 \Rightarrow 2$ and $2\Rightarrow4$ threads; first is major; second
        only minor
    \item Flamegraph still shows much cycles running against hurdles::Barrier::wait $\Rightarrow$
        this leaves room for improvement
    \item Flamegraph also shows that $\approx 75\%$ of all cycles is related to actual work (not
        just waiting).
        In particular to update\_mechanics functions.
\end{itemize}

>>>>

The developed methods are independent of how we assign parameters within our model.
This means we can opt to incorporate intracellular variability by specifying parameters on a
per-cell basis.
Or we can employ shared parameters between all agents for processes which can be expected to behave
identical within margin of error.
Furthermore, mixed approaches where some parameters are assigned globally and others individually
can be realized as well.

For the presented model a list of parameters can be seen in table~\ref{table:list-of-parameters} but
the following calculations are equally true for any model with $N_p$ number of parameters.
Consider the case where we filter out a subset $N_g\leq N_p$ of parameters which will be treated
globally, thus with equal numerical value for every cell.
This means that $N_p - N_g$ parameters will be assigned individually for each of $N_c$ number of
cells.
The total amount of parameters in the system is then given by
\begin{equation}
    N_\text{tot} = N_g + N_c (N_p - N_g).
\end{equation}
It is clear, that we should avoid, to needlessly treat parameters non-globally since this increases
the size of the parameter space considerably.
The volume of parameter space which needs to be sampled by optimization schemes scales exponentially
with the dimension ie. the total number of parameters $N_\text{tot}$.
Notice that the presented approach reflects our current work but can easily be adapted to other
scenarios where different strategies of parameters assignment are desired, e.g. when considering
two distinct species of cells which should be displayed by 2 parameters in total regardless of the
number of cells.

@@

We have already suspected (see Table~\ref{table:simulation-aspects}) that in order to properly
estimate parameters of our system, we need to account for natural cell-variability, thus assigning
parameters on a per-cell basis.

@@

\begin{itemize}
    \item Quadratic Scaling (see figure~\ref{fig:timings-crm_divide}) comes from more dense
        simulation
    \item Overall number of agents which can be simulated is satisfactory for current purposes
    \item Improvement from $1 \Rightarrow 2$ and $2\Rightarrow4$ threads; first is major; second
        only minor
    \item Flamegraph still shows much cycles running against hurdles::Barrier::wait $\Rightarrow$
        this leaves room for improvement
    \item Flamegraph also shows that $\approx 75\%$ of all cycles is related to actual work (not
        just waiting).
        In particular to update\_mechanics functions.
\end{itemize}
